---
title: "Example forest plots"
description: |
  Looking at different options to produce forest plots.
author:
  - name: Mark Baillie 
    url: https://bailliem.github.io/visR-gists/
    affiliation: visR
    affiliation_url: https://github.com/openpharma/visR
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    self_contained: FALSE
    toc: TRUE
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r setupR, results=FALSE, warning=FALSE}
library(rmarkdown)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(here)
library(patchwork)

theme_set(theme_minimal(base_size = 8))
#theme_set(theme_linedraw(base_size = 8))
```


# Purpose

Use ggplot and patchwork to combine plot and table objects in to forest plots, meta-analysis, grables, etc. 


# Layout

A forest plot is a table with a visual component i.e. a dotplot with uncertainty displaying summary/aggregated statistics (e.g. mean and 95% CI, median and 2.5th and 97.5th percentile of a posterior distribution) per row. The table may be a colection of columns or nested columns displaying the label per row, summary statistics and inferential statistics. 

See here for an example of layouts:

* https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/layouttypes.html
* https://cran.r-project.org/web/packages/RBesT/vignettes/customizing_plots.html
* https://github.com/malcolmbarrett/tidymeta
* http://rpkgs.datanovia.com/survminer/reference/ggforest.html

Can have nested columns (block by treatment)
Can have nexted rows (categories by factor or subgroup, study, etc. )

Table is a coordinate system columns and rows which are cells. 
Need to arrange / process data then have numbers or text then render as text or symbols. Text is symbols. Could do both.

Render using ggplot, format tables, whatever is there

What does our package handle. The transform step?

Broom - tidy estimates? 

Is this more than one package?
Is this a wrapper only?


# TODO / NOTES

Check tidymeta which renders table and plot in ggplot
https://github.com/malcolmbarrett/tidymeta

gt may this option:

* https://github.com/rstudio/gt/issues/152
* https://community.rstudio.com/t/report-with-compound-values-and-charts-together/22518/15
* https://github.com/rstudio/gt/pull/155



https://cran.r-project.org/web/packages/meta/

# Read data

Data is a summary of medical history measured at *diagnosis*:

* AFib 
* MI
* CHF
* Diabetes 
* Crohn's disease	


```{r}
crs <- read_csv(here("data","med_hist.csv"))
```

# Derive meta-data

Calculate the number of patients.

```{r}
bigN <- nrow(crs)
print(bigN)
```

# Aggregate

Process steps:

* Select comorbidities
* Transform data frame from wide to long
* Transform and spell out factors for plotting
* Plot counts by comorbidity 

TODO: Change dataset examples. 

```{r}
a_crs <- 
  crs %>%
  select(id, chf, afib, diabetes, mi, copd, crohn, ulcercol, ibd) %>%
  pivot_longer(-id, names_to = "comorb", values_to = "value") %>%
  mutate(
    comorb = case_when(
      comorb == "chf" ~ "Congestive heart failure",
      comorb == "afib" ~ "Atrial fibrillation",
      comorb == "diabetes" ~ "Diabetes",
      comorb == "copd" ~ "COPD",
      comorb == "mi" ~ "Myocardial infarction",
      comorb == "crohn" ~ "Crohn's disease",
      comorb == "ulcercol" ~ "Ulcerative colitis",
      comorb == "ibd" ~ "Inflammatory bowel disease"
    ),
    value = case_when(value == "0" ~ "no",
                      value == "1" ~ "yes")
  ) %>%
  group_by(comorb, value) %>%
  summarise(n = n()) %>%
  mutate(
    inc = n / bigN,
    perc = n / bigN * 100,
    plot_lab = paste0('(', n, ', ', round(perc, digits = 1), '%)'),
    axis_lab = paste0(value, ' ', '(', n, ', ', round(perc, digits = 1), '%)')
  ) 
```


# Plot table

Create two example table columns. 

```{r}

col_1 <- 
  a_crs %>%
  filter(value == "yes") %>%
  ggplot(aes(reorder(axis_lab, inc), 1, label = comorb), hjust = 0) +
  geom_text() +
  coord_flip() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank()
  )



col_1
```



```{r}
col_2 <- 
  a_crs %>%
  filter(value == "yes") %>%
  ggplot(aes(reorder(axis_lab, inc), 1, label = axis_lab), hjust = 0) +
  geom_text() +
  coord_flip() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank()
  )

col_2
```

# Plot numbers


This is an example of a dot plot with uncertainty. 

```{r}
col_3 <- 
  a_crs %>%
  filter(value == "yes") %>%
  ggplot(aes(reorder(axis_lab, inc), inc)) +
  geom_point() + 
  geom_linerange(aes(ymin = inc - 0.1, ymax = inc + 0.1), alpha = 0.2) +
  coord_flip() +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank()
  ) +
  ylab("Measurement [units]")

col_3
```



# Combine using patchwork

This is messy but illustrates this can be achieved. 

We are combining columns to form a dummy forest plot. 

```{r}
plot <- col_1 + col_2 + col_3
plot
```

Using this approach, will rows always been aligned?


Annotated version 
```{r}

plot + plot_annotation(
  title = 'Forest plot',
  subtitle = '3 plots combined into one',
  caption = 'Source data is from XYZ.'
)

```


# Session info


```{r}
sessionInfo()
```

# To read 
https://www.littlemissdata.com/blog/prettytables
https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/layouttypes.html

