---
title: "Example survival plots"
description: |
  Looking at different options to produce survival (time to event) plot including survminer, ggfortify and broom.
author:
  - name: Mark Baillie 
    url: https://bailliem.github.io/visR-gists/
    affiliation: visR
    affiliation_url: https://github.com/openpharma/visR
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
  toc: true
---

```{r setup, include=FALSE, results=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(rmarkdown)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(here)
library(patchwork)
library(survminer)
library(ggfortify)
library(survival)
library(broom)

#theme_set(theme_minimal(base_size = 8))
theme_set(theme_linedraw(base_size = 8))

```


# Purpose

To explore different options to produce a survival plot. 

# Example 

A typical example survival model with a binary stratum. 

```{r}
fit <- survfit(Surv(time, status) ~ sex, data = lung)
fit
```


# Survminer

Plot survival curves by strata using survminer. 

```{r}
ggsurvplot(fit, data = lung)
```


```{r}
ggsurvplot(
   fit,                     # survfit object with calculated statistics.
   data = lung,  # data used to fit survival curves. 
   risk.table = TRUE,         # show risk table.
   pval = TRUE,               # show p-value of log-rank test.
   conf.int = TRUE,           # show confidence intervals for 
                              # point estimaes of survival curves.
   xlim = c(0,1000),          # present narrower X axis, but not affect
                              # survival estimates.
   break.time.by = 500,       # break X axis in time intervals by 500.
   ggtheme = theme_minimal(), # customize plot and risk table with a theme.
   risk.table.y.text.col = T, # colour risk table text annotations.
   risk.table.y.text = FALSE # show bars instead of names in text annotations
                             # in legend of risk table
)
```



## GGfortify


```{r}
autoplot(fit)
```

## Broom 

First tidy the survival object. 

```{r}
td <- tidy(fit)
td %>% glimpse()
```

Then plot with ggplot. 

```{r}
surv_plot <- td %>% 
  ggplot(aes(time, estimate, group = strata)) + 
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)

surv_plot
```


Plot risk table 

```{r}
surv_tab <-
  td %>% 
  mutate(
    row = if_else(strata == "sex=1", 1, 0)  
  ) %>%
  ggplot(aes(x = time, y = row , label = n.risk)) + 
  geom_text() 

surv_tab
```

```{r}
surv_plot / surv_tab
```

How to recreate the risk table in surv_miner. 

Need to:

* within strata
* go through the sequence of time points to plot
* find last time point relative to each break
* round up if not exactly measured at that time point i.e. time = 8 becomes 10


```{r}

td

# this is the sequence
breaks <- c(0, 200, 400, 600, 800, 1000, 1200)

# filtering on the sequence
td %>%
  group_by(strata) %>%
  arrange(time) %>%
  filter(time < breaks) %>%
  slice(n()) %>%
  ungroup()
```

```{r}
sumfit <- sumfit <- summary(fit, times=c(0, 250, 500, 750, 1000, 1250))
n_risk <- data_frame('Arm' = sumfit$strata,
                     'Time' =  sumfit$time,
                     'Nrisk' =   sumfit$n.risk)

labels <- data_frame ('Arm' = "Time",
                      'Time' =  sumfit$time,
                      'Nrisk' =   sumfit$time)


n_risk
```

```{r}
surv_tab <-
  n_risk %>% 
  mutate(
    row = factor(if_else(Arm == "sex=1", 1, 0))  
  ) %>%
  ggplot(aes(x = Time, y = row , label = Nrisk)) + 
  geom_text() 
  

surv_tab
```

```{r}
surv_plot / surv_tab
```



## No confidence intervals 

```{r}
td %>% 
  ggplot(aes(time, estimate, group = strata, colour = strata)) + 
  geom_line() 

```

## Confidence interval

```{r}
td %>% 
  ggplot(aes(time, estimate, group = strata)) + 
  geom_line() +
  geom_linerange(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) 

```

## Faceted

```{r}
td %>% 
  ggplot(aes(time, estimate, group = strata)) + 
  geom_line() +
  geom_linerange(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  facet_wrap( ~ strata)

```

## With censoring


```{r}
td %>% 
  ggplot(aes(time, estimate, group = strata)) + 
  geom_line() +
  geom_point(data = td %>% filter(n.censor > 0), aes(time, estimate), alpha = 0.5)

```


## With annotations 


```{r}
library(ggrepel)
td %>% 
  ggplot(aes(time, estimate, group = strata)) + 
  geom_line() +
  geom_text_repel(data = td %>% group_by(strata) %>% slice(n()),
    aes(label = strata),
    size = 4.5,
    point.padding = .2,
    box.padding = .3,
    force = 1,
    min.segment.length = 0
  )

```


# Further reading

https://socviz.co/modeling.html
