---
title: "Revising a dynamite plot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(patchwork)


```

Have a clear purpose (worksheet)
Identify a project you are working on where you need to communicate in a data-driven way. Reflect upon and fill out the following.
What is the purpose of the project?
•  What is the main objective of the project?
•  List the (scientific) question(s) the project is trying to answer. Try to be specific.
•  What is the key evidence that is available to answer the question?
Who is your audience?
•  List the primary groups or individuals you will be communicating to.
•  If you had to narrow that to a single person, who would that be?
•  What does your audience care about?
•  What action does your audience need to take?
What is the importance of this project?
•  What are the benefits if your audience acts in the way that you want them to?
•  What are the risks if they do not?
What is the key message (the so what?)
•  Write out in a single sentence the key message.


```{r}
```


```{r}
set.seed(1971)

N <- 100
trt <- rep(c(0, 1), each=N/2)
aval <- rnorm(n = N, mean = 30 + trt * 1, sd = 5 + trt * 0.5)

dat <- tibble::tibble(trt = trt,
              aval = aval)


```

```{r}
mm1 <- function(...) {
  mean_cl_normal(...,mult=1)
}
mm2 <- function(...) {
  mean_cl_normal(...,mult=2)
}
```


```{r}
dat %>%
  ggplot(aes(x=factor(trt),y=aval))+
  stat_summary(fun.data=mean_cl_normal,geom="errorbar",width=0.5) +
  stat_summary(fun.data=mean_cl_normal,geom="bar") 
```

```{r}
g1 <-
  dat %>%
  ggplot(aes(x = factor(trt), y = aval)) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "errorbar",
    width = 0.5
  ) +
  stat_summary(fun.y = mean, geom = "bar") 
```


```{r}
g2 <-
  dat %>%
  ggplot(aes(x = factor(trt), y = aval)) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "errorbar",
    width = 0.5
  ) +
  stat_summary(fun.y = mean, geom = "bar") +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  ) 

```

```{r}

g3 <- 
  dat %>%
  ggplot(aes(x = factor(trt), y = aval)) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "errorbar",
    width = 0.5
  ) +
  stat_summary(fun.y = mean, geom = "bar") +
  theme_minimal(base_size = 14)+
  theme(legend.position="none",
        panel.grid.minor=element_blank(),
        panel.grid.major=element_line(size = 0.5, color = "gray96"),
        panel.background=element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank()
  )

```



```{r}
g4 <- 
  dat %>%
  ggplot(aes(x = factor(trt), y = aval)) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "errorbar",
    width = 0.5
  ) +
  stat_summary(fun.y = mean, geom = "bar") +
  xlab("Treatment") + ylab("Effect") + 
  theme_minimal(base_size = 14) +
  theme(legend.position="none",
        panel.grid.minor=element_blank(),
        panel.grid.major=element_line(size = 0.5, color = "gray96"),
        panel.background=element_blank()
  ) 

```



```{r}
g5 <-
  dat %>%
  ggplot(aes(x = factor(trt), y = aval)) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "linerange",
    size = 0.7
  ) +
  stat_summary(fun.y = mean, geom = "point") +
  xlab("Treatment") + ylab("Effect") +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(size = 0.5, color = "gray96"),
    panel.background = element_blank()
  )
```


```{r}
g6 <-
  dat %>%
  ggplot(aes(x = factor(trt), y = aval)) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "linerange",
    size = 1.1,
    color = "red",
    position = position_dodge(width = 0)
  ) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 2),
    geom = "linerange",
    size = 0.7
  ) +
  stat_summary(fun.y = mean, geom = "point") +
  xlab("Treatment") + ylab("Effect") +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(size = 0.5, color = "gray96"),
    panel.background = element_blank()
  )
```



```{r}
(g1 + g2 + g3) / (g4 + g5 + g6)
```
```{r}

dat %>%
  ggplot(aes(x = factor(trt), y = aval)) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "linerange",
    size = 1.5,
    color = "red",
    position = position_dodge(width = 0)
  ) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 2),
    geom = "linerange",
    size = 0.7
  ) +
  geom_jitter(aes(colour = factor(trt)), width = 0.1, alpha = 0.4) +
  xlab("Treatment") + ylab("Effect") +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(size = 0.5, color = "gray96"),
    panel.background = element_blank()
  )
```



```{r}

dat %>%
  ggplot(aes(x = factor(trt), y = aval)) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "linerange",
    size = 1.1,
    color = "red",
    position = position_dodge(width = 0)
  ) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 2),
    geom = "linerange",
    size = 0.7
  ) +
  geom_point(aes(colour = factor(trt)), alpha = 0.4) +
  xlab("Treatment") + ylab("Effect") +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(size = 0.5, color = "gray96"),
    panel.background = element_blank()
  )
```


