---
title: "Faceted bar chart"
date: "`r Sys.Date()`"
output:
    bookdown::html_document2:
        toc: true
        toc_float:
            collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r setupR, results=FALSE, warning=FALSE}
library(rmarkdown)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(here)
```

# About 

A summary of medical history measured at *diagnosis*:

* AFib, 
* MI
* CHF
* Diabetes 
* Crohn's disease	

Plot of the distribution of patients with specific comorbidities measured at diagnosis. 

Process steps:

* Select comorbidities
* Transform data frame from wide to long
* Transform and spell out factors for plotting
* Plot counts by comorbidity 

# Read data

```{r}
crs <- read_csv(here("data-raw","med_hist.csv"))
```

# Derive meta-data

Calculate the number of patients.

```{r}
bigN <- nrow(crs)
print(bigN)
```

# Aggregate


```{r}
a_crs <- 
  crs %>%
  select(id, chf, afib, diabetes, mi, copd, crohn, ulcercol, ibd) %>%
  pivot_longer(-id, names_to = "comorb", values_to = "value") %>%
  mutate(
    comorb = case_when(
      comorb == "chf" ~ "Congestive heart failure",
      comorb == "afib" ~ "Atrial fibrillation",
      comorb == "diabetes" ~ "Diabetes",
      comorb == "copd" ~ "COPD",
      comorb == "mi" ~ "Myocardial infarction",
      comorb == "crohn" ~ "Crohn's disease",
      comorb == "ulcercol" ~ "Ulcerative colitis",
      comorb == "ibd" ~ "Inflammatory bowel disease"
    ),
    value = case_when(value == "0" ~ "no",
                      value == "1" ~ "yes")
  ) %>%
  group_by(comorb, value) %>%
  summarise(n = n()) %>%
  mutate(
    inc = n / bigN,
    perc = n / bigN * 100,
    plot_lab = paste0('(', n, ', ', round(perc, digits = 1), '%)'),
    axis_lab = paste0(value, ' ', '(', n, ', ', round(perc, digits = 1), '%)')
  ) 
```


# Plot

```{r}
a_crs %>% 
  ggplot(aes(axis_lab, n)) +
  geom_col(width = 0.6,
           alpha = 0.15,
           fill = "black") +
  coord_flip() +
  facet_wrap(~ comorb, ncol = 2, scales = "free_y") +
  ggtitle("Number and percentage of patients reporting\n a comorbidity at diagnosis") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.title = element_blank()
  )
```


# Session info


```{r}
sessionInfo()
```